---
title: 'Project'
author: 
- "Ashish Ashish"
- "Arash Ahmadi"
date: 'November 7th, 2024'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
require(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(gtsummary)


library(readr)
olympics <- read_csv(here::here("data/olympics.csv"))
```


I added two new columns:

1. toal_gdp, the dataset already has gdp per capita

2. medal_pect, which is total medals divided by total athletes

```{r}
# make the chr to factor
olympics <- olympics %>% 
  mutate(country = factor(country),
         country_code = factor(country_code),
         region = factor(region))

olympics <- olympics %>% 
  mutate(total_gdp = gdp * population,
         medal_pct = total / athletes)

summary(olympics)
```
```{r}
olympics %>% 
  ggplot(aes(x=reorder(region, gdp, FUN = median), y=gdp, col=region)) +
  geom_boxplot() + 
  labs(title = "GDP per Capita vs Region", 
       x = "Region",
       y = "Total GDP")
ggsave("chart1.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
```

```{r}
olympics %>% 
  ggplot(aes(x=reorder(region, total, FUN = median), y=total, col=region)) +
  geom_boxplot() + 
  labs(title = "Total Medals vs Region", 
       x = "Region",
       y = "Total Medals")
ggsave("chart1.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
```


```{r}
olympics %>% 
  ggplot(aes(x=total_gdp, y=total, col=population)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(
    title = "Total Medals"
  )
ggsave("chart1.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
olympics %>% 
  ggplot(aes(x=total_gdp, y=gold, col=population)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(
    title = "Total Gold Medals"
  )
ggsave("chart2.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
olympics %>% 
  ggplot(aes(x=total_gdp, y=silver, col=population)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(
    title = "Total Silver Medals"
  )
ggsave("chart3.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
olympics %>% 
  ggplot(aes(x=total_gdp, y=bronze, col=population)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(
    title = "Total Bronze Medals"
  )
ggsave("chart4.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
```

**Check Conditions**:

Linearity
```{r}
ggplot(olympics, aes(x=total_gdp, y=total)) +
  geom_point()
ggsave("chart1.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
```


Independence: 

Normality of the residuals
```{r}
# Run SLR model
mod <- lm(total ~ total_gdp, data = olympics)
summary(mod)

# Check Normality Assumption
qqnorm(resid(mod))
qqline(resid(mod), col = "red")
ggsave("chart1.png", plot = last_plot(), width = 8, height = 6, dpi = 300)

hist(resid(mod))
ggsave("chart2.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
```

Equal (constant) Variance of the residuals
```{r}
plot(resid(mod) ~ fitted(mod), main = "Residuals vs. Fitted")  
abline(h = 0, col = "red") 
```

```{r}
hist(olympics$total_gdp)
```


I checked the assumptions for total gdp, and most of them were not meet, therefore, we transformed the gdp by "log function" and retest all the assumptions.

```{r}
olympics <- olympics %>% 
  mutate(log_total_gdp = log(total_gdp))

hist(olympics$log_total_gdp)

```


Check Assumptions
```{r}
ggplot(olympics, aes(x = log_total_gdp, y = log(total))) +
  geom_point()

# Run model with transformed x
mod_new <- lm(total ~ log_total_gdp, data = olympics)

# Check Normality Assumption
qqnorm(resid(mod_new))
qqline(resid(mod_new), col = "red")

# Check Equal (Constant) Variance Assumption
plot(resid(mod_new) ~ fitted(mod_new), main = "Residuals vs. Fitted") 
abline(h = 0, col = "red")

summary(mod_new)

```


Refrences:

chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://repository.gatech.edu/server/api/core/bitstreams/1aa2b537-c3de-4177-8295-3fcd3a03a965/content

chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://faculty.tuck.dartmouth.edu/images/uploads/faculty/andrew-bernard/olymp60restat_finaljournalversion.pdf